---
title: "ExData_Plotting1_Rcode"
author: "Xu Xu"
date: "8/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the data

```{r download zip into R and then unzip the file}
zip.url <- "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip"
dir <- getwd()
zip.file <- "data.zip"
zip.combine <- as.character(paste(dir, zip.file, sep = ";"))

download.file(zip.url, destfile = zip.combine)

unzip(zip.file)
```

```{r read in the "txt" data file}
df <- read.table("household_power_consumption.txt", sep = ";", na.strings ="?", header = TRUE)
head(df)
```

```{r Working with Date and Time}

df$date2 <- strptime(df$Date, "%d/%m/%Y")

class(df$date2)

#subsetting data from the dates 2007-02-01 and 2007-02-02

df_sub <- df[df$date2 >= "2007-02-01" & df$date2 <= "2007-02-02", ]

```

```{r}
df_sub$datetime <- paste(df_sub$Date, df_sub$Time)
df_sub$datetime2 <- strptime(df_sub$datetime, "%d/%m/%Y %H:%M:%S")
```


# Making Plots

## Plot 1 is a histogram of Global Active Power

- Note: The bars are red

```{r Plot 1}
png("plot1.png", width = 480, height= 480)
hist(df_sub$Global_active_power, main = "Global Active Power", xlab = "Global Active Power (kilowatts)", col = "red", yaxt="n") # yaxt = "n" to "Turn off y-axis"
# customized y-axis
axis(2, seq(0,1200, 200 ))

# save plot to file

dev.off()

```
## Plot 2 

Looking at the x-axis makes me realize that I need to convert the Date to the Day of the Week. (I knew how to do this before but I forgot! So I goggled.But in the end I did not use this new variable. Hmm.... )

```{r Date to "Day of the Week"}
df_sub$weekday <- weekdays(df_sub$date2) #conversion
table(df_sub$weekday) #quick look at the days
#Now I freak out... Only Thursday and Friday in my subset?
```

```{r Plot 2}
png("plot2.png", width = 480, height= 480)
plot(df_sub$Global_active_power,type="l",xaxt="n", ylab = "Global Active Power (kilowatts)", xlab = "") # x-axis being drawn usign xaxt="n"
# We want to plot a custom x-axis
# Let's put a date lable in three difference locations
locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))

# Customize the x-axis
axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))

# Save plot to file

dev.off()

```

## Plot 3 

Three layers in one plot. I am more familer with this using ggplot. How to do it in base R?

```{r Plot 3}
png("plot3.png", width = 480, height= 480)
# Create the first line
plot(df_sub$Sub_metering_1,type="l",xaxt="n", ylab = "Energy sub metering", xlab = "")
# Add a second line
lines(df_sub$Sub_metering_2,type="l",col = "red")
# Add a third line
lines(df_sub$Sub_metering_3,type="l",col = "blue")
# Add a legend to the plot
legend("topright", legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"), lty =1, col = c("black", "red", "blue"))

locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))

axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))

# Save plot to file
#dev.copy(png,'plot3.png') ## Notice that the legend got cut-off a bit. Need to readjust (see line 1)
dev.off()



```

## Plot 4

Building multiple plots (2 by 2 in this case). They share the same x- variable but have different y-variable. 

The the sub-plot in the lower left is plot 3 (three layers.)

```{r Plot 4}
png("plot4.png", width = 480, height= 480)
#set up the plot window
par(mfrow = c(2, 2)) #fill by row first

# Subplot 1 

plot(df_sub$Global_active_power, type="l",xaxt="n", ylab = "Global Active Power", xlab = "")
locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))
axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))

# Subplot 2
plot(df_sub$Voltage,type="l", ylab = "Voltage" , xaxt="n", xlab = "", sub = "datetime")
locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))
axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))


# Subplot 3
plot(df_sub$Sub_metering_1,type="l",xaxt="n", ylab = "Energy sub metering", xlab = "")
  ## Add a second line
lines(df_sub$Sub_metering_2,type="l",col = "red")
  ## Add a third line
lines(df_sub$Sub_metering_3,type="l",col = "blue")
  ## Add a legend to the plot
legend("topright", legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"), lty =1, col = c("black", "red", "blue"))

locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))
axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))
 
# Subplot 4

with(df_sub,plot(Global_reactive_power, type="l",xaxt="n", ylab = "Global_reactive_power", xlab = "", sub = "datetime"))

locations = floor(seq(from=0,to=nrow(df_sub),by=nrow(df_sub)/2))
axis(side = 1, at = locations, labels=c("Thu", "Fri", "Sat"))

# Save plot to file
dev.copy(png,'plot4.png')
dev.off()
```

